---
title: "Take-home Exercise 02"
format:
  html:
    code-fold: true
    code-summary: "Show the code"

execute: 
  eval: true
  echo: true
  freeze: true
  warning: false
  message: false
date: "`r Sys.Date()`"
---

# 1.0 Setting the Scene

Dengue Hemorrhagic Fever (in short dengue fever) is one of the most widespread mosquito-borne diseases in the most tropical and subtropical regions. It is an acute disease caused by dengue virus infection which is transmitted by female Aedes aegypti and Aedes albopictus mosquitoes. In 2015, Taiwan had recorded the most severe dengue fever outbreak with more than 43,000 dengue cases and 228 deaths. Since then, the annual reported dengue fever cases were maintained at the level of not more than 200 cases. However, in 2023, Taiwan recorded 26703 dengue fever cases.

# 2.0 Objectives

Discover:

-   if the distribution of dengue fever outbreak at Tainan City, Taiwan are independent from space and space and time.

-   If the outbreak is indeed spatial and spatio-temporal dependent, then, you would like to detect where are the clusters and outliers, and the emerging hot spot/cold spot areas.

# 3.0 The Task

-   Using appropriate function of **sf** and **tidyverse**, preparing the following geospatial data layer:

    -   a study area layer in sf polygon features. It must be at village level and confined to the D01, D02, D04, D06, D07, D08, D32 and D39 counties of Tainan City, Taiwan.

    -   a dengue fever layer within the study area in sf point features. The dengue fever cases should be confined to epidemiology week 31-50, 2023.

    -   a derived dengue fever layer in [spacetime s3 class of sfdep](https://sfdep.josiahparry.com/articles/spacetime-s3). It should contain, among many other useful information, a data field showing number of dengue fever cases by village and by epidemiology week.

-   Using the extracted data, perform global spatial autocorrelation analysis by using [sfdep methods](https://is415-gaa-tskam.netlify.app/in-class_ex/in-class_ex05/in-class_ex05-glsa).

-   Using the extracted data, perform local spatial autocorrelation analysis by using [sfdep methods](https://r4gdsa.netlify.app/chap10.html).

-   Using the extracted data, perform emerging hotspot analysis by using [sfdep methods](https://is415-gaa-tskam.netlify.app/in-class_ex/in-class_ex05/in-class_ex05-ehsa).

-   Describe the spatial patterns revealed by the analysis above.

# 4.0 The Data

Two data sets

-   TAIWAN_VILLAGE_2020, a geospatial data of village boundary of Taiwan. It is in ESRI shapefile format. The data is in Taiwan Geographic Coordinate System. (Source: [Historical map data of the village boundary: TWD97 longitude and latitude](https://data.gov.tw/en/datasets/130549))

-   Dengue_Daily.csv, an aspatial data of reported dengue cases in Taiwan since 1998. (Source: [Dengue Daily Confirmed Cases Since 1998](https://data.cdc.gov.tw/en/dataset/dengue-daily-determined-cases-1998). Below are selected fields that are useful for this study:

    -   發病日: Onset date

    -   最小統計區中心點X: x-coordinate

    -   最小統計區中心點Y: y-coordinate

# 5.0 Importing and uploading packages

```{r}
pacman::p_load(sf, tidyverse, sfdep, lubridate, spdep, tmap, dplyr)

```

# 6.0 Importing the data into R environment

## 6.1 Importing aspatial data

```{r}
dengue <- read_csv("data/aspatial/Dengue_Daily.csv")
```

## 6.2 Importing geospatial data

```{r}
tainan_village <- st_read(dsn = "data/geospatial", layer="TAINAN_VILLAGE")


```

# 7.0 Data Wrangling

```{r}
filter_dengue <- dengue[c(1,10,11)]
```

```{r}
filter_tainan_village <- tainan_village %>% filter(TOWNID %in% c("D01", "D02", "D04", "D06", "D07", "D08", "D32", "D39"))
```

```{r}
plot(st_geometry(filter_tainan_village))
```

```{r}
basemap <- tm_shape(filter_tainan_village) +
  tm_polygons() +
  tm_text("TOWNID", size=0.5)
tmap_arrange(basemap, ncol=1)
```

```{r}
```

