---
title: "In Class Exercise 6 -Geographical Segmentation with Spatially Constrained Clustering Techniques"
format:
  html:
    code-fold: true
    code-summary: "Show the code"

execute: 
  eval: true
  echo: true
  warning: false
date: "`r Sys.Date()`"
---

# 1.0 Installing and loading R packages

```{r}
pacman::p_load(spdep, tmap, sf, ClustGeo, 
               ggpubr, cluster, factoextra, NbClust,
               heatmaply, corrplot, psych, tidyverse, GGally, sp)
```

# 2.0 Data Import and Preparation

```{r}
shan_sf <- st_read(dsn = "data/geospatial", 
                   layer = "myanmar_township_boundaries") %>%
  filter(ST %in% c("Shan (East)", "Shan (North)", "Shan (South)")) %>%
  select(c(2:7))
```

```{r}
class(shan_sf)
```


view the content of the newly created shan_sf simple features data.frame

```{r}
shan_sf
```

Notice that sf.data.frame is conformed to Hardy Wickham’s tidy framework.

Since shan_sf is conformed to tidy framework, we can also glimpse() to reveal the data type of it’s fields.

```{r}
glimpse(shan_sf)
```

## 2.1 Importing aspatial data into R environment

```{r}
ict <- read_csv ("data/aspatial/Shan-ICT.csv")
```

Summary statistics of ict data.frame.

```{r}
summary(ict)
```

## 2.2 Derive new variable using dplyr package

The unit of measurement of the values are number of household. Using these values directly will be bias by the underlying total number of households. In general, the townships with relatively higher total number of households will also have higher number of households owning radio, TV, etc.

In order to overcome this problem, we will derive the penetration rate of each ICT variable

```{r}
ict_derived <- ict %>%
  mutate(`RADIO_PR` = `Radio`/`Total households`*1000) %>%
  mutate(`TV_PR` = `Television`/`Total households`*1000) %>%
  mutate(`LLPHONE_PR` = `Land line phone`/`Total households`*1000) %>%
  mutate(`MPHONE_PR` = `Mobile phone`/`Total households`*1000) %>%
  mutate(`COMPUTER_PR` = `Computer`/`Total households`*1000) %>%
  mutate(`INTERNET_PR` = `Internet at home`/`Total households`*1000) %>%
  rename(`DT_PCODE` =`District Pcode`,`DT`=`District Name`,
         `TS_PCODE`=`Township Pcode`, `TS`=`Township Name`,
         `TT_HOUSEHOLDS`=`Total households`,
         `RADIO`=`Radio`, `TV`=`Television`, 
         `LLPHONE`=`Land line phone`, `MPHONE`=`Mobile phone`,
         `COMPUTER`=`Computer`, `INTERNET`=`Internet at home`) 
```

Review the summary statistics of the newly derived penetration rates

```{r}
summary(ict_derived)
```

Notice that six new fields have been added into the data.frame. They are RADIO_PR, TV_PR, LLPHONE_PR, MPHONE_PR, COMPUTER_PR, and INTERNET_PR.

# 3.0 Exploratory Data Analysis

Plot the distribution of the variables (i.e. Number of households with radio) by using appropriate Exploratory Data Analysis (EDA)

Histogram is useful to identify the overall distribution of the data values (i.e. left skew, right skew or normal distribution)

```{r}
ggplot(data=ict_derived, 
       aes(x=`RADIO`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")
```

Boxplot is useful to detect if there are outliers.

```{r}
ggplot(data=ict_derived, 
       aes(x=`RADIO`)) +
  geom_boxplot(color="black", 
               fill="light blue")
```

Plotting the distribution of the newly derived variables (i.e. Radio penetration rate)

```{r}
ggplot(data=ict_derived, 
       aes(x=`RADIO_PR`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")
```

```{r}
ggplot(data=ict_derived, 
       aes(x=`RADIO_PR`)) +
  geom_boxplot(color="black", 
               fill="light blue")
```

Multiple histograms are plotted to reveal the distribution of the selected variables in the ict_derived data.frame.

Consist of two main parts. First, we will create the individual histograms

```{r}
radio <- ggplot(data=ict_derived, 
             aes(x= `RADIO_PR`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

tv <- ggplot(data=ict_derived, 
             aes(x= `TV_PR`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

llphone <- ggplot(data=ict_derived, 
             aes(x= `LLPHONE_PR`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

mphone <- ggplot(data=ict_derived, 
             aes(x= `MPHONE_PR`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

computer <- ggplot(data=ict_derived, 
             aes(x= `COMPUTER_PR`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

internet <- ggplot(data=ict_derived, 
             aes(x= `INTERNET_PR`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

```

Next, the ggarrange() function of ggpubr package is used to group these histograms together.

```{r}
ggarrange(radio, tv, llphone, mphone, computer, internet, 
          ncol = 3, 
          nrow = 2)
```

## 3.1 EDA usingn choropleth map

### 3.1.1 Joining geospatial data with aspatial data

Need to combine both the geospatial data object (i.e. shan_sf) and aspatial data.frame object (i.e. ict_derived) into one. This will be performed by using the left_join function of dplyr package. The shan_sf simple feature data.frame will be used as the base data object and the ict_derived data.frame will be used as the join table.

The unique identifier used to join both data objects is TS_PCODE.

Geometry file should be first. 
st_join, both file must be sf file.
```{r}
#| eval: False
shan_sf <- left_join(shan_sf, 
                     ict_derived, by=c("TS_PCODE"="TS_PCODE"))
  
write_rds(shan_sf,"data/rds/shan_sf.rds")
```

The message above shows that TS_CODE field is the common field used to perform the left-join.

It is important to note that there is no new output data been created. Instead, the data fields from ict_derived data frame are now updated into the data frame of shan_sf.

```{r}
shan_sf <- read_rds("data/rds/shan_sf.rds")
```

### 3.1.2 Prepare the choropleth map

Quick look at the distribution of Radio penetration rate of Shan State at township level, a choropleth map will be prepared.

```{r}
qtm(shan_sf, "RADIO_PR")
```

In order to reveal the distribution shown in the choropleth map above are bias to the underlying total number of households at the townships, we will create two choropleth maps, one for the total number of households (i.e. TT_HOUSEHOLDS.map) and one for the total number of household with Radio (RADIO.map)

```{r}
TT_HOUSEHOLDS.map <- tm_shape(shan_sf) + 
  tm_fill(col = "TT_HOUSEHOLDS",
          n = 5,
          style = "jenks", 
          title = "Total households") + 
  tm_borders(alpha = 0.5) 

RADIO.map <- tm_shape(shan_sf) + 
  tm_fill(col = "RADIO",
          n = 5,
          style = "jenks",
          title = "Number Radio ") + 
  tm_borders(alpha = 0.5) 

tmap_arrange(TT_HOUSEHOLDS.map, RADIO.map,
             asp=NA, ncol=2)
```

**Notice that the choropleth maps above clearly show that townships with relatively larger number ot households are also showing relatively higher number of radio ownership.**

Plot the choropleth maps showing the dsitribution of total number of households and Radio penetration rate.

```{r}
tm_shape(shan_sf) +
    tm_polygons(c("TT_HOUSEHOLDS", "RADIO_PR"),
                style="jenks") +
    tm_facets(sync = TRUE, ncol = 2) +
  tm_legend(legend.position = c("right", "bottom"))+
  tm_layout(outer.margins=0, asp=0)
```

# 4.0 Correlation Analysis

It is important for us to ensure that the cluster variables are not highly correlated.

In this section, you will learn how to use corrplot.mixed() function of corrplot package to visualise and analyse the correlation of the input variables.

```{r}
cluster_vars.cor = cor(ict_derived[,12:17])
corrplot.mixed(cluster_vars.cor,
         lower = "ellipse", 
               upper = "number",
               tl.pos = "lt",
               diag = "l",
               tl.col = "black")
```

The correlation plot above shows that COMPUTER_PR and INTERNET_PR are highly correlated. This suggest that only one of them should be used in the cluster analysis instead of both.

# 5.0 Hierarchical Cluster Analysis

The analysis consists of four major steps:

## 5.1 Extracting clustering variables

Extract the clustering variables from the shan_sf simple feature object into data.frame.

```{r}
cluster_vars <- shan_sf %>%
  st_set_geometry(NULL) %>%
  select("TS.x", "RADIO_PR", "TV_PR", "LLPHONE_PR", "MPHONE_PR", "COMPUTER_PR")
head(cluster_vars,10)
```

Notice that the final clustering variables list does not include variable INTERNET_PR because it is highly correlated with variable COMPUTER_PR.

Next, change the rows by township name instead of row number

```{r}
row.names(cluster_vars) <- cluster_vars$"TS.x"
head(cluster_vars,10)
```

Notice that the row number has been replaced into the township name.

Now, we will delete the TS.x field.

```{r}
shan_ict <- select(cluster_vars, c(2:6))
head(shan_ict, 10)
```

## 5.2 Data Standardization

In general, multiple variables will be used in cluster analysis. It is not unusual their values range are different. In order to avoid the cluster analysis result is baised to clustering variables with large values, it is useful to standardise the input variables before performing cluster analysis.

## 5.3 Min-Max standaraisation

Normalize() of heatmaply package is used to stadardisation the clustering variables by using Min-Max method. The summary() is then used to display the summary statistics of the standardised clustering variables.

```{r}
shan_ict.std <- normalize(shan_ict)
summary(shan_ict.std)
```

Notice that the values range of the Min-max standardised clustering variables are 0-1 now.

## 5.4 Z-score standardization

Z-score standardisation can be performed easily by using scale() of Base R. The code chunk below will be used to stadardisation the clustering variables by using Z-score method.

```{r}
shan_ict.z <- scale(shan_ict)
describe(shan_ict.z)
```

Notice the mean and standard deviation of the Z-score standardised clustering variables are 0 and 1 respectively.

Note: describe() of psych package is used here instead of summary() of Base R because the earlier provides standard deviation.

**Warning: Z-score standardisation method should only be used if we would assume all variables come from some normal distribution.**

## 5.5 Visualising the standardised clustering variables

Beside reviewing the summary statistics of the standardised clustering variables, it is also a good practice to visualise their distribution graphical.

```{r}
r <- ggplot(data=ict_derived, 
             aes(x= `RADIO_PR`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue") +
  ggtitle("Raw values without standardisation")

shan_ict_s_df <- as.data.frame(shan_ict.std)
s <- ggplot(data=shan_ict_s_df, 
       aes(x=`RADIO_PR`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue") +
  ggtitle("Min-Max Standardisation")

shan_ict_z_df <- as.data.frame(shan_ict.z)
z <- ggplot(data=shan_ict_z_df, 
       aes(x=`RADIO_PR`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue") +
  ggtitle("Z-score Standardisation")

ggarrange(r, s, z,
          ncol = 3,
          nrow = 1)
```

```{r}
r <- ggplot(data=ict_derived, 
             aes(x= `RADIO_PR`)) +
  geom_density(color="black",
               fill="light blue") +
  ggtitle("Raw values without standardisation")

shan_ict_s_df <- as.data.frame(shan_ict.std)
s <- ggplot(data=shan_ict_s_df, 
       aes(x=`RADIO_PR`)) +
  geom_density(color="black",
               fill="light blue") +
  ggtitle("Min-Max Standardisation")

shan_ict_z_df <- as.data.frame(shan_ict.z)
z <- ggplot(data=shan_ict_z_df, 
       aes(x=`RADIO_PR`)) +
  geom_density(color="black",
               fill="light blue") +
  ggtitle("Z-score Standardisation")

ggarrange(r, s, z,
          ncol = 3,
          nrow = 1)
```

## 5.6 Computing proximity matrix

In R, many packages provide functions to calculate distance matrix. We will compute the proximity matrix by using dist() of R.

dist() supports six distance proximity calculations, they are: euclidean, maximum, manhattan, canberra, binary and minkowski. The default is euclidean proximity matrix.

Compute the proximity matrix using euclidean method.

```{r}
proxmat <- dist(shan_ict, method = 'euclidean')
```

List the content of proxmat for visual inspection.

```{r}
proxmat
```

## 5.7 Computing hierarchical clustering

In R, there are several packages provide hierarchical clustering function. In this hands-on exercise, hclust() of R stats will be used.

hclust() employed agglomeration method to compute the cluster. Eight clustering algorithms are supported, they are: ward.D, ward.D2, single, complete, average(UPGMA), mcquitty(WPGMA), median(WPGMC) and centroid(UPGMC).

Performs hierarchical cluster analysis using ward.D method. The hierarchical clustering output is stored in an object of class hclust which describes the tree produced by the clustering process.

```{r}
hclust_ward <- hclust(proxmat, method = 'ward.D')
```

Plot the tree by using plot() of R Graphics.

```{r}
plot(hclust_ward, cex = 0.6)
```

## 5.8 Selecting the optimal clustering algorithm

One of the challenge in performing hierarchical clustering is to identify stronger clustering structures. The issue can be solved by using use agnes() function of cluster package. It functions like hclus(), however, with the agnes() function you can also get the agglomerative coefficient, which measures the amount of clustering structure found (values closer to 1 suggest strong clustering structure).

Compute the agglomerative coefficients of all hierarchical clustering algorithms.

```{r}
m <- c( "average", "single", "complete", "ward")
names(m) <- c( "average", "single", "complete", "ward")

ac <- function(x) {
  agnes(shan_ict, method = x)$ac
}

map_dbl(m, ac)
```

With reference to the output above, we can see that Ward’s method provides the strongest clustering structure among the four methods assessed. Hence, in the subsequent analysis, only Ward’s method will be used.

## 5.9 Determining Optimal clusters

Another technical challenge face by data analyst in performing clustering analysis is to determine the optimal clusters to retain.

There are three commonly used methods to determine the optimal clusters, they are:

Elbow Method Average Silhouette Method Gap Statistic Method

## 5.9.1 Gap Statistic Method

The gap statistic compares the total within intra-cluster variation for different values of k with their expected values under null reference distribution of the data. The estimate of the optimal clusters will be value that maximize the gap statistic (i.e., that yields the largest gap statistic). This means that the clustering structure is far away from the random uniform distribution of points.

To compute the gap statistic, clusGap() of cluster package will be used.

```{r}
set.seed(12345)
gap_stat <- clusGap(shan_ict, 
                    FUN = hcut, 
                    nstart = 25, 
                    K.max = 10, 
                    B = 50)
print(gap_stat, method = "firstmax")
```

Also note that the hcut function used is from factoextra package.

Next,visualise the plot by using fviz_gap_stat() of factoextra package.

use 6 cluster base on the graph below.

```{r}
fviz_gap_stat(gap_stat)
```

With reference to the gap statistic graph above, the recommended number of cluster to retain is 1. However, it is not logical to retain only one cluster. By examine the gap statistic graph, the 6-cluster gives the largest gap statistic and should be the next best cluster to pick.

Note: In addition to these commonly used approaches, the NbClust package, published by Charrad et al., 2014, provides 30 indices for determining the relevant number of clusters and proposes to users the best clustering scheme from the different results obtained by varying all combinations of number of clusters, distance measures, and clustering methods.

## 5.10 Interpreting the dendograms

In the dendrogram displayed above, each leaf corresponds to one observation. As we move up the tree, observations that are similar to each other are combined into branches, which are themselves fused at a higher height.

The height of the fusion, provided on the vertical axis, indicates the (dis)similarity between two observations. The higher the height of the fusion, the less similar the observations are. Note that, conclusions about the proximity of two observations can be drawn only based on the height where branches containing those two observations first are fused. We cannot use the proximity of two observations along the horizontal axis as a criteria of their similarity.

It’s also possible to draw the dendrogram with a border around the selected clusters by using rect.hclust() of R stats. The argument border is used to specify the border colors for the rectangles.

```{r}
plot(hclust_ward, cex = 0.6)
rect.hclust(hclust_ward, 
            k = 6, 
            border = 2:5)
```

## 5.11 Visually-driven hierarchical cluster analysis

In this section, we will learn how to perform visually-driven hiearchical clustering analysis by using heatmaply package.

With heatmaply, we are able to build both highly interactive cluster heatmap or static cluster heatmap.

### 5.11.1 Transforming the data frame into a matrix

The data was loaded into a data frame, but it has to be a data matrix to make your heatmap.

Transforming shan_ict data frame into a data matrix.

```{r}
shan_ict_mat <- data.matrix(shan_ict)
```

### 5.11.2 Plotting interactive cluster heatmap using heatmaply()

The heatmaply() of heatmaply package is used to build an interactive cluster heatmap.

```{r}
heatmaply(normalize(shan_ict_mat),
          Colv=NA,
          dist_method = "euclidean",
          hclust_method = "ward.D",
          seriate = "OLO",
          colors = Blues,
          k_row = 6,
          margins = c(NA,200,60,NA),
          fontsize_row = 4,
          fontsize_col = 5,
          main="Geographic Segmentation of Shan State by ICT indicators",
          xlab = "ICT Indicators",
          ylab = "Townships of Shan State"
          )
```

## 5.12 Mapping the clusters formed

With closed examination of the dendragram above, we have decided to retain six clusters.

cutree() of R Base will be used in the code chunk below to derive a 6-cluster model.

```{r}
groups <- as.factor(cutree(hclust_ward, k=6))
```

The output is called groups. It is a list object.

In order to visualise the clusters, the groups object need to be appended onto shan_sf simple feature object.

The code chunk below form the join in three steps:

-   the groups list object will be converted into a matrix;

-   cbind() is used to append groups matrix onto shan_sf to produce an output simple feature object called shan_sf_cluster; and

-   rename of dplyr package is used to rename as.matrix.groups field as CLUSTER.


cbind is column binding. No sorting, if not cbind is not allow

```{r}
shan_sf_cluster <- cbind(shan_sf, as.matrix(groups)) %>%
  rename(`CLUSTER`=`as.matrix.groups.`)
```

Next, qtm() of tmap package is used to plot the choropleth map showing the cluster formed.

```{r}
qtm(shan_sf_cluster, "CLUSTER")
```

The choropleth map above reveals the clusters are very fragmented. The is one of the major limitation when non-spatial clustering algorithm such as hierarchical cluster analysis method is used.

# 6.0 Spatially Constrained Clustering: SKATER approach

Derive spatially constrained cluster by using skater() method of spdep package.

## 6.1 Converting into SpatialPloygonsDataFrame

First,convert shan_sf into SpatialPolygonsDataFrame. This is because SKATER function only support sp objects such as SpatialPolygonDataFrame.

Use as_Spatial() of sf package to convert shan_sf into a SpatialPolygonDataFrame called shan_sp.

```{r}
shan_sp <- as_Spatial(shan_sf)
```

## 6.2 Computing Neighbour

List Next, poly2nd() of spdep package will be used to compute the neighbours list from polygon list.

```{r}
shan.nb <- poly2nb(shan_sp)
summary(shan.nb)
```

Plot the neighbours list on shan_sp. Since we now can plot the community area boundaries as well, we plot this graph on top of the map. The first plot command gives the boundaries. This is followed by the plot of the neighbor list object, with coordinates applied to the original SpatialPolygonDataFrame (Shan state township boundaries) to extract the centroids of the polygons. These are used as the nodes for the graph representation. We also set the color to blue and specify add=TRUE to plot the network on top of the boundaries.

```{r}
plot(shan_sp, 
     border=grey(.5))
plot(shan.nb, 
     coordinates(shan_sp), 
     col="blue", 
     add=TRUE)
```

```{r}
plot(st_geometry(shan_sf), 
     border=grey(.5))
pts <- st_coordinates(st_centroid(shan_sf))
plot(shan.nb, 
     pts, 
     col="blue", 
     add=TRUE)
```

Note that if you plot the network first and then the boundaries, some of the areas will be clipped. This is because the plotting area is determined by the characteristics of the first plot. In this example, because the boundary map extends further than the graph, we plot it first.

## 6.3 Computing minimum spanning tree

### 6.3.1 Calculating edge cost

Next, nbcosts() of spdep package is used to compute the cost of each edge. It is the distance between it nodes. This function compute this distance using a data.frame with observations vector in each node.

```{r}
lcosts <- nbcosts(shan.nb, shan_ict)
```

For each observation, this gives the pairwise dissimilarity between its values on the five variables and the values for the neighbouring observation (from the neighbour list). Basically, this is the notion of a generalised weight for a spatial weights matrix.

Next, We will incorporate these costs into a weights object in the same way as we did in the calculation of inverse of distance weights. In other words, we convert the neighbour list to a list weights object by specifying the just computed lcosts as the weights.

In order to achieve this, nb2listw() of spdep package is used as shown in the code chunk below.

Note that we specify the style as B to make sure the cost values are not row-standardised.

```{r}
shan.w <- nb2listw(shan.nb, 
                   lcosts, 
                   style="B")
summary(shan.w)
```

## 6.4 Computing minimum spanning tree

The minimum spanning tree is computed by mean of the mstree() of spdep package

```{r}
shan.mst <- mstree(shan.w)
```

After computing the MST, we can check its class and dimension.

```{r}
class(shan.mst)
```

```{r}
dim(shan.mst)
```

Note that the dimension is 54 and not 55. This is because the minimum spanning tree consists on n-1 edges (links) in order to traverse all the nodes.

We can display the content of shan.mst by using head()

```{r}
head(shan.mst)
```

The plot method for the MST include a way to show the observation numbers of the nodes in addition to the edge. As before, we plot this together with the township boundaries. We can see how the initial neighbour list is simplified to just one edge connecting each of the nodes, while passing through all the nodes.

```{r}
plot(shan_sp, border=gray(.5))
plot.mst(shan.mst, 
         coordinates(shan_sp), 
         col="blue", 
         cex.lab=0.7, 
         cex.circles=0.005, 
         add=TRUE)
```

## 6.5 Computing spatially constrained clusters using SKATERS method

Compute the spatially constrained cluster using skater() of spdep package.

```{r}
clust6 <- spdep::skater(edges = shan.mst[,1:2], 
                 data = shan_ict, 
                 method = "euclidean", 
                 ncuts = 5)
```

The skater() takes three mandatory arguments: - the first two columns of the MST matrix (i.e. not the cost), - the data matrix (to update the costs as units are being grouped), and - the number of cuts. Note: It is set to one less than the number of clusters. So, the value specified is not the number of clusters, but the number of cuts in the graph, one less than the number of clusters.

The result of the skater() is an object of class skater. We can examine its contents

```{r}
str(clust6)
```

The most interesting component of this list structure is the groups vector containing the labels of the cluster to which each observation belongs (as before, the label itself is arbitary). This is followed by a detailed summary for each of the clusters in the edges.groups list. Sum of squares measures are given as ssto for the total and ssw to show the effect of each of the cuts on the overall criterion.

check the cluster assignment by using the conde chunk below.

```{r}
ccs6 <- clust6$groups
ccs6
```

We can find out how many observations are in each cluster by means of the table command. Parenthetially, we can also find this as the dimension of each vector in the lists contained in edges.groups. For example, the first list has node with dimension 12, which is also the number of observations in the first cluster.

```{r}
table(ccs6)
```

Lastly, we can also plot the pruned tree that shows the five clusters on top of the townshop area.

```{r}
plot(shan_sp, border=gray(.5))
plot(clust6, 
     coordinates(shan_sp), 
     cex.lab=.7,
     groups.colors=c("red","green","blue", "brown", "pink"),
     cex.circles=0.005, 
     add=TRUE)
```

## 6.6 Visualizing the clusters in choropleth map

Plot the newly derived clusters by using SKATER method.

```{r}
groups_mat <- as.matrix(clust6$groups)
shan_sf_spatialcluster <- cbind(shan_sf_cluster, as.factor(groups_mat)) %>%
  rename(`SP_CLUSTER`=`as.factor.groups_mat.`)
qtm(shan_sf_spatialcluster, "SP_CLUSTER")
```

For easy comparison, it will be better to place both the hierarchical clustering and spatially constrained hierarchical clustering maps next to each other.

```{r}
hclust.map <- qtm(shan_sf_cluster,
                  "CLUSTER") + 
  tm_borders(alpha = 0.5) 

shclust.map <- qtm(shan_sf_spatialcluster,
                   "SP_CLUSTER") + 
  tm_borders(alpha = 0.5) 

tmap_arrange(hclust.map, shclust.map,
             asp=NA, ncol=2)
```

# 7.0 Spatially Constrained Clustering: ClustGeo Method

Using functions provided by ClustGeo package to perform non-spatially constrained hierarchical cluster analysis and spatially constrained cluster analysis.

## 7.1 A short Note about ClustGeo pacakge

ClustGeo package is an R package specially designed to support the need of performing spatially constrained cluster analysis. More specifically, it provides a Ward-like hierarchical clustering algorithm called hclustgeo() including spatial/geographical constraints.

In the nutshell, the algorithm uses two dissimilarity matrices D0 and D1 along with a mixing parameter alpha, whereby the value of alpha must be a real number between \[0, 1\]. D0 can be non-Euclidean and the weights of the observations can be non-uniform. It gives the dissimilarities in the attribute/clustering variable space. D1, on the other hand, gives the dissimilarities in the constraint space. The criterion minimised at each stage is a convex combination of the homogeneity criterion calculated with D0 and the homogeneity criterion calculated with D1.

The idea is then to determine a value of alpha which increases the spatial contiguity without deteriorating too much the quality of the solution based on the variables of interest. This need is supported by a function called choicealpha().

## 7.2 Ward-like hierarchical clustering: ClustGeo

ClustGeo package provides function called hclustgeo() to perform a typical Ward-like hierarchical clustering just like hclust() you learned in previous section.

To perform non-spatially constrained hierarchical clustering, we only need to provide the function a dissimilarity matrix as shown in the code chunk below.

```{r}
nongeo_cluster <- hclustgeo(proxmat)
plot(nongeo_cluster, cex = 0.5)
rect.hclust(nongeo_cluster, 
            k = 6, 
            border = 2:5)
```

Note that the dissimilarity matrix must be an object of class dist, i.e. an object obtained with the function dist(). For sample code chunk, please refer to 5.7.6 Computing proximity matrix

### 7.2.1 Mapping the clusters formed

Similarly, we can plot the clusters on a categorical area shaded map by using the steps we learned in 5.7.12 Mapping the clusters formed.

```{r}
groups <- as.factor(cutree(nongeo_cluster, k=6))
```

```{r}
shan_sf_ngeo_cluster <- cbind(shan_sf, as.matrix(groups)) %>%
  rename(`CLUSTER` = `as.matrix.groups.`)
```

```{r}
qtm(shan_sf_ngeo_cluster, "CLUSTER")
```

## 7.3 Spatially Constrained Hierarchical Clustering

Before we can performed spatially constrained hierarchical clustering, a spatial distance matrix will be derived by using st_distance() of sf package.

```{r}
dist <- st_distance(shan_sf, shan_sf)
distmat <- as.dist(dist)
```

Notice that as.dist() is used to convert the data frame into matrix.

Next, choicealpha() will be used to determine a suitable value for the mixing parameter alpha as shown in the code chunk below.

```{r}
cr <- choicealpha(proxmat, distmat, range.alpha = seq(0, 1, 0.1), K=6, graph = TRUE)
```

With reference to the graphs above, alpha = 0.3 will be used.

```{r}
clustG <- hclustgeo(proxmat, distmat, alpha = 0.3)
```

Next, cutree() is used to derive the cluster object.

```{r}
groups <- as.factor(cutree(clustG, k=6))
```

Join back the group list with shan_sf polygon feature data frame.

```{r}
shan_sf_Gcluster <- cbind(shan_sf, as.matrix(groups)) %>%
  rename(`CLUSTER` = `as.matrix.groups.`)
```

Plot the map of the newly delineated spatially constrained clusters.

```{r}
qtm(shan_sf_Gcluster, "CLUSTER")
```

# 8.0 Visual Interpretation of Clusters

## 8.1 Visualizing individual clustering variable

Used to reveal the distribution of a clustering variable (i.e RADIO_PR) by cluster.

```{r}
ggplot(data = shan_sf_ngeo_cluster,
       aes(x = CLUSTER, y = RADIO_PR)) +
  geom_boxplot()
```

The boxplot reveals Cluster 3 displays the highest mean Radio Ownership Per Thousand Household. This is followed by Cluster 2, 1, 4, 6 and 5.

## 8.2 Multivariate Visualization

Past studies shown that parallel coordinate plot can be used to reveal clustering variables by cluster very effectively. In the code chunk below, ggparcoord() of GGally package

```{r}
ggparcoord(data = shan_sf_ngeo_cluster, 
           columns = c(17:21), 
           scale = "globalminmax",
           alphaLines = 0.2,
           boxplot = TRUE, 
           title = "Multiple Parallel Coordinates Plots of ICT Variables by Cluster") +
  facet_grid(~ CLUSTER) + 
  theme(axis.text.x = element_text(angle = 30))
```

The parallel coordinate plot above reveals that households in Cluster 4 townships tend to own the highest number of TV and mobile-phone. On the other hand, households in Cluster 5 tends to own the lowest of all the five ICT.

Note that the scale argument of ggparcoor() provide several methods to scale the clustering variables. They are:

std: univariately, subtract mean and divide by standard deviation. robust: univariately, subtract median and divide by median absolute deviation. uniminmax: univariately, scale so the minimum of the variable is zero, and the maximum is one. globalminmax: no scaling is done; the range of the graphs is defined by the global minimum and the global maximum. center: use uniminmax to standardize vertical height, then center each variable at a value specified by the scaleSummary param. centerObs: use uniminmax to standardize vertical height, then center each variable at the value of the observation specified by the centerObsID param There is no one best scaling method to use. You should explore them and select the one that best meet your analysis need.

Last but not least, we can also compute the summary statistics such as mean, median, sd, etc to complement the visual interpretation.

Group_by() and summarise() of dplyr are used to derive mean values of the clustering variables.

```{r}
shan_sf_ngeo_cluster %>% 
  st_set_geometry(NULL) %>%
  group_by(CLUSTER) %>%
  summarise(mean_RADIO_PR = mean(RADIO_PR),
            mean_TV_PR = mean(TV_PR),
            mean_LLPHONE_PR = mean(LLPHONE_PR),
            mean_MPHONE_PR = mean(MPHONE_PR),
            mean_COMPUTER_PR = mean(COMPUTER_PR))
```
